---
title: "Model: yield_tha ~ surrounding landscape + intrinsic variaton + field design"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(refund)
library(mgcv)
```

# Contents 

(1) Model Structure
(1) Refund::prf Introduction to the function used
(2) Data Exploration 
(3) Model run
(4) Validation and Verification

# Model Structure 

*yield_tha ~ landscape heterogeneity (swf) + intrinsic variaton + field design*

Because the surrounding-landscape predictor is a curve rather than a
single number, the response is modelled with a scalar-on-function
regression (`refund::pfr`), which is a generalized additive model
(`mgcv`) extended to accept functional covariates.




This document analyses crop yield (`yield_tha`) in an agroforestry
system as a function of three groups of predictors:

1. **surrounding landscape** - the proportion of small woody features
   (SWF) around the field, expressed as a *function* of distance from
   the field boundary (100-1000 m, 100 m steps);
2. **intrinsic variation** - the distance from the sampled plot to the
   tree strip within the field;
3. **field design** - fixed characteristics of each field/year
   (climate, soil, landscape configuration, crop, tree age, planting
   year).

Because the surrounding-landscape predictor is a curve rather than a
single number, the response is modelled with a scalar-on-function
regression (`refund::pfr`), which is a generalized additive model
(`mgcv`) extended to accept functional covariates.

The data-exploration section below follows the structure recommended by
Zuur et al. 2010:
- outlier
- distribution/homogeneity of the response
- zero-inflation
- collinearity 
- relationships 
- independence of observations.

```{r data}
df <- read.csv("C:/Users/Elizaveta/OneDrive - Universität Bayreuth/Dokumente/MasterThesis/MA_RProject/data/AnalysisData/20260819_AFswf.csv")

df <- df |>
  mutate(year_AFplanting = case_when(
    field == "Dornburg"      ~ 2007,
    field == "Forst"         ~ 2010,
    field == "Gladbacherhof" ~ 2020,
    field == "IhingerHof"    ~ 2008,
    field == "Mariensee"     ~ 2008,
    field == "Reiffenhausen" ~ 2011,
    field == "Vechta"        ~ 2019,
    field == "Wendhausen"    ~ 2008
  )) |> 
  mutate(AFage = year - year_AFplanting)

fert_raw <- gsub("[\u2013\u2212]", "-", df$fertilization_rate_kg_n_p_k_ha_1_year_1)
df$fert_N <- as.numeric(sub("^([0-9.]+).*", "\\1", fert_raw))
df$fert_N[fert_raw == "no fertilization"] <- 0

# one row per plot-year observation, for exploration of scalar variables
base_explore <- df |>
  distinct(id, field, year, yield_tha, distance_to_tree_strip,
           crop_unified, area, fert_N, temp_C_mean, sun_MJ_m2_mean,
           precip_mm_sum, l_shdi, l_contag, l_ed, l_ai,
           mean_slope, min_slope, max_slope, treeage,
           year_AFplanting, prop_swf_within)
```

# Data Exploration 

## Distribution of the response variable: yield tha

```{r distribution of yield}

ggplot(df, aes(x = seq_along(yield_tha), y = yield_tha)) +
  geom_point(shape = 1, alpha = 0.5) +
  theme_bw() +
  labs(
    x = "Observation",
    y = "Yield (t/ha)"
  )

ggplot(df, aes(x = seq_along(yield_tha), y = yield_tha, color = crop_unified)) +
  geom_point(shape = 1, alpha = 0.5) +
  theme_bw() +
  labs(
    x = "Observation",
    y = "Yield (t/ha)"
  )

ggplot(df, aes(x = seq_along(yield_tha), y = yield_tha, color = distance_to_tree_strip)) +
  geom_point(shape = 1, alpha = 0.5) +
  theme_bw() +
  labs(
    x = "Observation",
    y = "Yield (t/ha)"
  )


```


## Outliers

Cleveland dotplots for every continuous variable (observation order on
the y-axis, value on the x-axis) make isolated extreme values easy to
spot without imposing a distributional assumption.

```{r outliers, fig.height=8, fig.width=7}
num_vars <- c("yield_tha", "distance_to_tree_strip", "fert_N",
              "temp_C_mean", "sun_MJ_m2_mean", "precip_mm_sum",
              "l_shdi", "l_contag", "l_ed", "l_ai",
              "mean_slope", "treeage", "prop_swf_within")

dotplot_data <- base_explore |>
  select(all_of(num_vars)) |>
  mutate(row = row_number()) |>
  pivot_longer(-row, names_to = "variable", values_to = "value")

ggplot(dotplot_data, aes(x = value, y = row)) +
  geom_point(shape = 1, alpha = 0.5) +
  facet_wrap(~variable, scales = "free_x") +
  theme_bw() +
  labs(y = "observation order", x = NULL,
       title = "Cleveland dotplots for outlier screening")
```

Flag anything that sits noticeably apart from the bulk of the points in
its panel; decide case by case whether it is a data-entry error (fix),
a genuine extreme value (keep, but check its influence on the fitted
model later), or an artefact of the extraction pipeline (e.g. a
boundary raster cell with no woody-cover pixels).

## Distribution / homogeneity of the response

```{r response-dist, fig.width=6, fig.height=4}
ggplot(base_explore, aes(x = yield_tha)) +
  geom_histogram(bins = 30, colour = "white") +
  theme_bw() +
  labs(title = "Distribution of yield_tha")

ggplot(base_explore, aes(x = crop_unified, y = yield_tha)) +
  geom_boxplot() +
  theme_bw() +
  labs(title = "yield_tha by crop", x = NULL) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

`yield_tha` differs strongly by crop (as the boxplot and the model's
parametric coefficients confirm), which is why `crop_unified` is kept
as a fixed factor rather than pooled. Within-crop spread should look
roughly comparable across fields/years for a Gaussian `pfr` to be an
appropriate choice; strong heteroscedasticity by crop would motivate a
`weights =` argument or a variance-modelling approach.

## Zero-inflation

```{r zero-check}
sapply(base_explore[num_vars], function(x) mean(x == 0, na.rm = TRUE))
```

None of the continuous predictors or the response should show a
disproportionate spike at zero (`fert_N` is expected to have a
genuine, meaningful cluster at 0 for "no fertilization" plots - that is
not zero-inflation in the statistical sense, just a real category, and
is handled by the smooth term being able to flatten out near 0).

## Collinearity among scalar predictors

```{r collinearity, fig.width=7, fig.height=6}
cor_mat <- base_explore |>
  select(all_of(num_vars[-1])) |>   # drop the response
  cor(use = "pairwise.complete.obs")

round(cor_mat, 2)
```

Pay particular attention to `area`, `mean_slope`, `min_slope`,
`max_slope`, `temp_C_mean`, `sun_MJ_m2_mean`, `precip_mm_sum`,
`l_shdi`, `l_contag`, `l_ed`, `l_ai` - with only 8 fields, several of
these are near-constant within a field and therefore highly correlated
with field identity itself (see Section 5.3, concurvity). `area` and
`mean_slope`/`min_slope`/`max_slope` are dropped from the model for
exactly this reason: each has a single value per field.

## Relationships between yield and candidate predictors

```{r relationships, fig.width=8, fig.height=6}
rel_data <- base_explore |>
  select(yield_tha, distance_to_tree_strip, fert_N, temp_C_mean,
         precip_mm_sum, l_shdi, treeage, crop_unified) |>
  pivot_longer(-c(yield_tha, crop_unified), names_to = "variable", values_to = "value")

ggplot(rel_data, aes(x = value, y = yield_tha)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "loess", se = FALSE) +
  facet_wrap(~ crop_unified + variable, scales = "free") +
  theme_bw()
```

```{r relationships-swf, fig.width=6, fig.height=4}
df |>
  select(id, distance, prop_swf) |>
  distinct() |>
  ggplot(aes(x = distance, y = prop_swf)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "loess", se = FALSE) +
  theme_bw() +
  labs(title = "prop_swf as a function of distance from the field boundary")
```

Loess curves that are clearly non-straight (as expected for
`distance_to_tree_strip` and for `prop_swf` vs. `distance`) are the
empirical justification for modelling these relationships with
penalized smooths / a functional term rather than linear terms.

## Independence of observations

```{r independence}
table(base_explore$field)
base_explore |> count(field, year) |> tidyr::pivot_wider(names_from = year, values_from = n)
```

Multiple plots per field and multiple years per field mean observations
are not independent at the field level (and possibly the plot level
across years). This nesting is why the model includes `s(field,
bs = "re")` rather than treating all 647 rows as independent
replicates; if a given `id`/plot recurs across years, consider adding a
plot-level random effect as well (see Section 5.5).

# Variable Overview

| Variable | Description | Unit | Role in model |
|---|---|---|---|
| `yield_tha` | Crop yield | t/ha | **Scalar response** |
| `W_swf` (built from `distance`, `prop_swf`) | Proportion of small woody features (SWF) in the annulus at each distance from the field boundary | proportion (0-1), function of distance (m) | **Functional predictor** (continuous), entered via `lf()` |
| `distance_to_tree_strip` | Distance from the sampled plot to the nearest agroforestry tree strip | m | **Continuous scalar predictor**, smooth term |
| `prop_swf_within` | Proportion of SWF inside the field's own circular footprint (radius = boundary_min) | proportion (0-1) | Continuous scalar predictor (field design context; consider as smooth or covariate) |
| `crop_unified` | Harmonised crop type grown that year | category (7 levels: grass, maize, potato, rapeseed, rye, wheat, ...) | **Categorical predictor**, parametric factor |
| `area` | Field area | m^2 (or ha, check source) | Field design - dropped (1 unique value per field, confounded with field RE) |
| `fert_N` | Nitrogen fertilization rate (parsed from the N-P-K string; "no fertilization" = 0) | kg N ha^-1 yr^-1 | **Continuous scalar predictor**, smooth term |
| `temp_C_mean` | Mean growing-season temperature | deg C | **Continuous scalar predictor**, smooth term |
| `sun_MJ_m2_mean` | Mean solar radiation | MJ/m^2 | **Continuous scalar predictor**, smooth term |
| `precip_mm_sum` | Total precipitation | mm | **Continuous scalar predictor**, smooth term |
| `l_shdi` | Shannon diversity index of the surrounding landscape (FRAGSTATS) | index | **Continuous scalar predictor**, smooth term |
| `l_contag` | Contagion index of the surrounding landscape | index | **Continuous scalar predictor**, smooth term |
| `l_ed` | Edge density of the surrounding landscape | m/ha | **Continuous scalar predictor**, smooth term |
| `l_ai` | Aggregation index of the surrounding landscape | index | **Continuous scalar predictor**, smooth term |
| `mean_slope`, `min_slope`, `max_slope` | Field topographic slope statistics | degrees | Field design - `mean_slope` dropped (1 value per field, confounded with field RE); `min_slope`/`max_slope` currently unused, candidates for the same treatment |
| `treeage` | Age of the agroforestry tree strip in the given year | years | **Continuous scalar predictor**, smooth term |
| `year_AFplanting` | Year the agroforestry system was planted at that field | year (categorical, 8 levels, 1 per field) | Field design - candidate for `factor()`/smooth; near-collinear with `field` |
| `field` | Field identifier | category (8 levels) | **Random-effect grouping factor**, `s(field, bs = "re")` |
| `id` | Unique plot-year observation identifier | - | Row/observation key, not a covariate |
| `year` | Calendar year of observation | year | Available for a year effect / repeated-measures structure, not currently in the model |
| `distance` | Distance from field boundary at which `prop_swf` was measured | m (100-1000, step 100) | Domain (`argvals`) of the functional predictor `W_swf`, not a covariate itself |

# Model Specification

## Building the functional covariate

`prop_swf` is recorded at 10 fixed distances (100-1000 m, 100 m steps)
per `id`. Pivoting to wide format gives one row per `id` and one column
per distance - this matrix is the discretised functional predictor
`W_swf(distance)`, and `grid` is its domain.

```{r build-W}
W_swf <- df |>
  dplyr::select(id, distance, swf_year, prop_swf) |>
  unique() |>
  pivot_wider(names_from = distance, values_from = prop_swf) |>
  arrange(id)

id_order <- W_swf$id
W_mat    <- as.matrix(W_swf[, -c(1, 2)])
grid     <- as.numeric(colnames(W_mat))

stopifnot(!anyNA(grid), length(grid) == ncol(W_mat))
```

## Building the scalar dataset

One row per `id`, aligned exactly to the row order of `W_mat`.

```{r build-base}
base <- df |>
  distinct(id, field, year, yield_tha, distance_to_tree_strip,
           crop_unified, area, fert_N, temp_C_mean, sun_MJ_m2_mean,
           precip_mm_sum, l_shdi, l_contag, l_ed, l_ai,
           mean_slope, min_slope, max_slope, treeage,
           year_AFplanting, prop_swf_within) |>
  filter(id %in% id_order) |>
  slice(match(id_order, id))

stopifnot(nrow(base) == nrow(W_mat))
base$W_swf <- W_mat

base$field         <- factor(base$field)
base$crop_unified  <- factor(base$crop_unified)
```

## Adaptive basis dimensions

`safe_k()` caps every smooth's basis dimension (`k`) at
`(number of unique values - 1)` of its own covariate, so that
`smooth.construct` can never fail regardless of how listwise deletion
(NAs anywhere in the formula) thins the fitting sample.

```{r safe-k}
safe_k <- function(x, k_wanted) {
  min(k_wanted, length(unique(x[!is.na(x)])) - 1)
}

k_dist  <- safe_k(base$distance_to_tree_strip, 6)
k_fertN <- safe_k(base$fert_N,                 5)
k_temp  <- safe_k(base$temp_C_mean,            5)
k_sun   <- safe_k(base$sun_MJ_m2_mean,         5)
k_prec  <- safe_k(base$precip_mm_sum,          5)
k_shdi  <- safe_k(base$l_shdi,                 5)
k_cont  <- safe_k(base$l_contag,               5)
k_ed    <- safe_k(base$l_ed,                   5)
k_ai    <- safe_k(base$l_ai,                   5)
k_tree  <- safe_k(base$treeage,                5)
```

## The model

```{r fit-model}
model <- pfr(
  yield_tha ~
    lf(W_swf, argvals = grid, bs = "ps", k = 15) +      # (1) surrounding landscape
    s(distance_to_tree_strip, bs = "tp", k = k_dist) +   # (2) intrinsic variation
    s(fert_N,         bs = "tp", k = k_fertN) +    # thin plate regression splines
    s(temp_C_mean,    bs = "tp", k = k_temp) +
    s(sun_MJ_m2_mean, bs = "tp", k = k_sun) +
    s(precip_mm_sum,  bs = "tp", k = k_prec) +
    s(l_shdi,         bs = "tp", k = k_shdi) +
    s(l_contag,       bs = "tp", k = k_cont) +
    s(l_ed,           bs = "tp", k = k_ed) +
    s(l_ai,           bs = "tp", k = k_ai) +
    s(treeage,        bs = "tp", k = k_tree) +
    crop_unified +
    s(field, bs = "re"), # random effects 
  data   = base,
  method = "REML"
)

summary(model)
```

Term-by-term rationale:

- `lf(W_swf, argvals = grid, bs = "ps", k = 15)` - the surrounding
  landscape term. Estimates a smooth coefficient function beta(distance)
  over the 100-1000 m domain using a penalized B-spline (`bs = "ps"`)
  basis; `k = 15` is generous relative to the 10 observed distances but
  is fine because the roughness penalty (chosen by REML) controls the
  effective wiggliness, not `k` itself.
- `s(distance_to_tree_strip, bs = "tp", k = k_dist)` - intrinsic
  variation. A thin-plate smooth lets yield respond nonlinearly to
  proximity to the tree strip, consistent with the competition-zone
  literature (Swieter et al. 2019).
  I THINK HERE ONE CAN JUST TAKE A LOG SMOOTH 
THERE IS NO INDICATION THAT THESE EFFECTS ARE NON LINEAR SO I WOULD TAKE THEM LINEARLY 
The block of `s(..., bs = "tp", k = ...)` terms for `fert_N`,
  climate normals, and landscape-configuration indices represents field
  design as smooth (not necessarily linear) scalar effects.
- `crop_unified` - parametric factor; crop identity has a large,
  roughly additive effect on yield and does not need smoothing.
- `s(field, bs = "re")` - random intercept for field, absorbing
  unmodelled, field-specific variation and acknowledging that
  observations within a field are not independent.

# Model Validation and Verification

## Basis dimension check (k-index)

```{r gam-check, fig.width=7, fig.height=7}
gam.check(model)
```

Look at the `k-index` column: a value well below 1 together with a
low p-value for a given smooth means `k` may be too small (the smooth
is being forced to be smoother than the data support) and should be
increased (subject to `safe_k`'s ceiling); values around/above 1 are
fine.

## Residual diagnostics

```{r residual-diagnostics, fig.width=8, fig.height=6}
res <- residuals(model)
fit <- fitted(model)

par(mfrow = c(2, 2))
plot(fit, res, main = "Residuals vs fitted", xlab = "Fitted", ylab = "Residuals")
abline(h = 0, col = "red")
hist(res, main = "Residual distribution", xlab = "Residuals")
qqnorm(res); qqline(res, col = "red")
# plot(base$distance_to_tree_strip, res,
#      main = "Residuals vs distance_to_tree_strip",
#      xlab = "distance_to_tree_strip", ylab = "Residuals")
# abline(h = 0, col = "red")
# par(mfrow = c(1, 1))
```

Check specifically for: (i) trumpet/fan shape in residuals-vs-fitted
(heteroscedasticity, possibly by crop - consider `s(fitted, by =
crop_unified)`-style variance modelling or a Gamma/log-link family if
present); (ii) heavy tails in the Q-Q plot (consider a robust family or
investigate the flagged outliers from Section 2.1); (iii) any remaining
pattern of residuals against `distance_to_tree_strip` or fitted values,
which would indicate the smooth is under- or over-fitting.

## Residuals by field and by year (checks the random-effect structure)

```{r residuals-by-group, fig.width=7, fig.height=4}
# resid_df <- base |> mutate(resid = res)
# 
# ggplot(resid_df, aes(x = field, y = resid)) +
#   geom_boxplot() +
#   geom_hline(yintercept = 0, colour = "red") +
#   theme_bw() +
#   labs(title = "Residuals by field") +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
# 
# ggplot(resid_df, aes(x = factor(year), y = resid)) +
#   geom_boxplot() +
#   geom_hline(yintercept = 0, colour = "red") +
#   theme_bw() +
#   labs(title = "Residuals by year", x = "year")
```

If residuals still show a systematic offset by field, `s(field,
bs = "re")` is not fully doing its job (often because of the
concurvity issue described next) and the fixed effects should be
simplified. A systematic pattern by year suggests adding a year term.

## Concurvity (the GAM analogue of collinearity)

```{r concurvity}
concurvity(model, full = TRUE)
concurvity(model, full = FALSE)
```

Given that several field design smooths are near-constant within a
field (Section 2.4), expect high concurvity between those terms and
`s(field, bs = "re")`. `full = FALSE` shows pairwise concurvity and
helps identify which specific pair of terms is problematic. If
concurvity between, e.g., `s(temp_C_mean)` and `s(field)` is high,
report this explicitly and consider the simplified model discussed in
`01_model_choice_rationale.txt` (Section 2) as a robustness check.

## Influence / leave-one-field-out sensitivity

```{r loo-field, eval=FALSE}
# refit dropping each field once, compare beta(distance) for W_swf
fields <- levels(base$field)
loo_coefs <- lapply(fields, function(f) {
  m <- update(model, data = base[base$field != f, ])
  coef(m, n = length(grid))
})
```

With only 8 fields, a leave-one-field-out refit is cheap and directly
answers "is the estimated surrounding-landscape effect (or the
distance-to-tree-strip effect) driven by a single field?" - important
given the concurvity concerns above. Compare the estimated beta(distance)
curve (and its confidence band) across the 8 refits, e.g. by overlaying
`coef(m, n = ...)` for each fold.

## Reported fit statistics

```{r fit-stats}
summary(model)$r.sq
summary(model)$dev.expl
AIC(model)
```

Report adjusted R^2, deviance explained, and REML/AIC alongside the
diagnostics above - a high R^2 (as in the current model, ~0.76) is not
by itself evidence that the model is correctly specified if concurvity
is high; the diagnostics in this section are what establishes that.



```{r PCA}
library(dplyr)
library(tidyr)
library(FactoMineR)
library(factoextra)
library(ggplot2)

W_swf <- df |>
  dplyr::select(id, distance, prop_swf) |>
  unique() |>
  filter(distance %in% c(100, 500, 1000)) |>
  pivot_wider(names_from = distance, values_from = prop_swf, names_prefix = "swf_") |>
  arrange(id)

pca_data <- df |>
  distinct(id, yield_tha, distance_to_tree_strip, crop_unified) |>
  inner_join(W_swf, by = "id") |>
  na.omit()

pca_res <- PCA(pca_data |> select(-id),
                quali.sup = which(names(pca_data |> select(-id)) == "crop_unified"),
                quanti.sup = which(names(pca_data |> select(-id)) == "yield_tha"),
                graph = FALSE)

get_eigenvalue(pca_res)

ind_coord <- as.data.frame(pca_res$ind$coord[, 1:2])
colnames(ind_coord) <- c("Dim1", "Dim2")
ind_coord$yield_scaled <- as.numeric(scale(pca_data$yield_tha))
ind_coord$crop_unified <- pca_data$crop_unified

var_coord <- as.data.frame(pca_res$var$coord[, 1:2])
colnames(var_coord) <- c("Dim1", "Dim2")
var_coord$label <- rownames(var_coord)

scale_f <- 0.8 * max(abs(ind_coord[, 1:2])) / max(abs(var_coord[, 1:2]))

ggplot() +
  geom_point(data = ind_coord,
             aes(x = Dim1, y = Dim2, color = yield_scaled, shape = crop_unified),
             size = 2.5) +
  geom_segment(data = var_coord,
               aes(x = 0, y = 0, xend = Dim1 * scale_f, yend = Dim2 * scale_f),
               arrow = arrow(length = unit(0.2, "cm")), color = "black") +
  geom_text(data = var_coord,
            aes(x = Dim1 * scale_f * 1.1, y = Dim2 * scale_f * 1.1, label = label),
            size = 3.5) +
  scale_color_gradient2(low = "#2c7fb8", mid = "#ffffbf", high = "#d7191c",
                         midpoint = 0) +
  theme_minimal() +
  labs(x = "Dim1", y = "Dim2", color = "yield_scaled", shape = "crop") +
  theme(panel.grid.minor = element_blank())

```


```{r}

swf_long <- df |>
  dplyr::select(id, field, distance, prop_swf) |>
  unique()

swf_wide <- swf_long |>
  tidyr::pivot_wider(names_from = distance, values_from = prop_swf) |>
  arrange(id)

grid <- as.numeric(names(swf_wide)[-c(1, 2)])
Y_swf <- as.matrix(swf_wide[, -c(1, 2)])
field_id <- df |> distinct(id, field) |> arrange(id) |> pull(field)

swf_model <- pffr(
  Y_swf ~ s(field_id, bs = "re"),
  yind = grid,
  data = list(Y_swf = Y_swf, field_id = factor(field_id)),
  family = binomial()
)

summary(swf_model)
plot(swf_model)

```
```{r}
fitted_swf <- fitted(swf_model)
rownames(fitted_swf) <- swf_wide$id

W_smooth <- fitted_swf[match(id_order, rownames(fitted_swf)), ]

base$W_swf <- W_smooth

model <- pfr(
  yield_tha ~
    lf(W_swf, argvals = grid, bs = "ps", k = 15) +
    s(distance_to_tree_strip, bs = "tp", k = k_dist) +
    s(fert_N,         bs = "tp", k = k_fertN) +
    s(temp_C_mean,    bs = "tp", k = k_temp) +
    s(sun_MJ_m2_mean, bs = "tp", k = k_sun) +
    s(precip_mm_sum,  bs = "tp", k = k_prec) +
    s(l_shdi,         bs = "tp", k = k_shdi) +
    s(l_contag,       bs = "tp", k = k_cont) +
    s(l_ed,           bs = "tp", k = k_ed) +
    s(l_ai,           bs = "tp", k = k_ai) +
    s(treeage,        bs = "tp", k = k_tree) +
    crop_unified +
    s(field, bs = "re"),
  data   = base,
  method = "REML"
)

summary(model)
```

