---
title: "Data Exploration Protocol"
subtitle: "Following Zuur, Ieno & Elphick (2010) — *Methods in Ecology and Evolution* 1: 3–14"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    number_sections: true
    df_print: paged
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  message   = FALSE,
  warning   = FALSE,
  fig.width = 8,
  fig.height = 5,
  dpi       = 150
)
```

```{r libraries}
library(tidyverse)   # wrangling + ggplot2
library(lattice)     # dotplots / coplots (Zuur's classic tool)
library(car)         # vif(), leveneTest(), scatterplotMatrix()
library(lme4)        # variance components (Step 8)
library(patchwork)   # combine ggplots side-by-side
```

```{r load-data}
# --- Load and type your data here -------------------------------------------
df <- readr::read_delim("../data/Koch25/yields_wintercrops.csv", delim = ";", locale = locale(decimal_mark = ","))
df <- janitor::clean_names(df)
#
df <- df |>
  mutate(
    block  = factor(block),
    id     = factor(id),
    year   = factor(year),
    p_dist = factor(p_dist)
  )
```

---

# Step 1 — Outliers in Y and X {.tabset}

> **Zuur p. 4:** Cleveland dotplots are preferred over boxplots for detecting
> outliers because they preserve observation order and reveal data-entry errors.

## Dotplot — `yield_wweight`

```{r step1-dotplot}
dotchart(
  df$yield_wweight,
  main  = "Cleveland dotplot — yield_wweight",
  xlab  = "yield_wweight",
  ylab  = "Observation order",
  pch   = 19,
  color = "steelblue"
)
```

## Boxplots by group

```{r step1-boxplots}
# p1 <- ggplot(df, aes(x = p_dist, y = yield_wweight)) +
#   geom_boxplot(fill = "steelblue", alpha = 0.7, outlier.colour = "red") +
#   labs(title = "yield_wweight by p_dist", x = "p_dist", y = "yield_wweight") +
#   theme_bw()

p2 <- ggplot(df, aes(x = factor(year), y = yield_wweight)) +
  geom_boxplot(fill = "darkorange", alpha = 0.7, outlier.colour = "red") +
  labs(title = "yield_wweight by year", x = "year", y = NULL) +
  theme_bw()

# p1 + p2
```

---

# Step 2 — Homogeneity of Variance {.tabset}

> **Zuur p. 6:** Heterogeneity of variance among groups violates OLS assumptions
> and is often missed without conditional plots. Levene's test provides a formal check.

## Plots

```{r step2-plots}
# p3 <- ggplot(df, aes(x = p_dist, y = yield_wweight, fill = p_dist)) +
#   geom_boxplot(alpha = 0.7, show.legend = FALSE) +
#   labs(title = "Variance by p_dist", x = "p_dist", y = "yield_wweight") +
#   theme_bw()

p4 <- ggplot(df, aes(x = factor(year), y = yield_wweight, fill = factor(year))) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  labs(title = "Variance by year", x = "year", y = NULL) +
  theme_bw()

p5 <- ggplot(df, aes(x = factor(block), y = yield_wweight, fill = factor(block))) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  labs(title = "Variance by block", x = "block", y = "yield_wweight") +
  theme_bw()

p4 + p5
```

## Levene's Test

```{r step2-levene}
# H0: variances are equal across groups
# leveneTest(yield_wweight ~ p_dist,        data = df)
leveneTest(yield_wweight ~ factor(year),  data = df)
```

---

# Step 3 — Normality of Y {.tabset}

> **Zuur p. 7:** Normality of the raw response and *within each group* matters.
> Histograms and QQ-plots are more informative than Shapiro–Wilk alone, which
> is overly sensitive at large *n*.

## Overall

```{r step3-overall, fig.width=9}
par(mfrow = c(1, 2))

hist(df$yield_wweight,
     breaks = 20,
     col    = "steelblue",
     border = "white",
     main   = "Histogram — yield_wweight",
     xlab   = "yield_wweight")

qqnorm(df$yield_wweight, main = "Normal QQ-plot — yield_wweight", pch = 19, col = "steelblue")
qqline(df$yield_wweight, col = "red", lwd = 2)

par(mfrow = c(1, 1))

shapiro.test(df$yield_wweight)
```

## By p_dist

```{r step3-pdist}
# Conditional histograms — the lattice way (Zuur Fig. 3)
histogram(~ yield_wweight | p_dist,
          data   = df,
          xlab   = "yield_wweight",
          main   = "yield_wweight distribution within each p_dist",
          layout = c(NA, 1),
          col    = "steelblue")
```

## By year

```{r step3-year}
histogram(~ yield_wweight | factor(year),
          data = df,
          xlab = "yield_wweight",
          main = "yield_wweight distribution within each year",
          col  = "darkorange")
```

## Shapiro–Wilk per group

```{r step3-shapiro}
df |>
  group_by(p_dist) |>
  summarise(
    n       = n(),
    W       = shapiro.test(yield_wweight)$statistic,
    p_value = shapiro.test(yield_wweight)$p.value
  )
```

---

# Step 4 — Zero Inflation {.tabset}

> **Zuur p. 8:** Excess zeros can invalidate Gaussian models and may indicate
> that a zero-inflated or Tweedie model is more appropriate.

## Summary

```{r step4-summary}
zero_pct <- mean(df$yield_wweight == 0) * 100
cat(sprintf("Zeros in yield_wweight: %.1f%% of %d observations\n",
            zero_pct, nrow(df)))

# Zeros by p_dist and year
df |>
  group_by(p_dist, year) |>
  summarise(
    n         = n(),
    n_zeros   = sum(yield_wweight == 0),
    zero_pct  = round(mean(yield_wweight == 0) * 100, 1),
    .groups   = "drop"
  )
```

## Plot

```{r step4-plot}
df |>
  mutate(is_zero = yield_wweight == 0) |>
  group_by(block, year) |>
  summarise(zero_pct = mean(is_zero) * 100, .groups = "drop") |>
  ggplot(aes(x = factor(year), y = zero_pct, fill = factor(block))) +
  geom_col(position = "dodge") +
  labs(title  = "Zero proportion by block × year",
       x      = "year",
       y      = "% zeros",
       fill   = "block") +
  theme_bw()
```

---

# Step 5 — Collinearity Among Covariates {.tabset}

> **Zuur p. 9:** Collinear predictors inflate standard errors and destabilise
> coefficient estimates. VIF > 3 warrants caution; VIF > 10 is serious.
> **Replace `covariate1`, `covariate2`, `covariate3` with your actual predictors.**

## Correlation matrix

```{r step5-cor}
covars <- df |> select(covariate1, covariate2, covariate3)  # ← adjust
cor(covars, use = "pairwise.complete.obs") |> round(2)
```

## Scatterplot matrix

```{r step5-scattermatrix, fig.height=6}
scatterplotMatrix(
  ~ covariate1 + covariate2 + covariate3,  # ← adjust
  data   = df,
  smooth = list(method = "loessLine"),
  main   = "Step 5 — Scatterplot matrix of covariates"
)
```

## VIF

```{r step5-vif}
m_vif <- lm(yield_wweight ~ covariate1 + covariate2 + covariate3, data = df)
vif(m_vif)
```

---

# Step 6 — Y–X Relationships {.tabset}

> **Zuur p. 10:** Scatterplots of Y against each covariate, conditioned on
> grouping factors, reveal non-linearity and guide link-function choice.

## By p_dist

```{r step6-pdist}
ggplot(df, aes(x = covariate1, y = yield_wweight, colour = p_dist)) +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE) +
  facet_wrap(~ p_dist) +
  labs(title = "yield_wweight ~ covariate1 | p_dist") +
  theme_bw() +
  theme(legend.position = "none")
```

## By year

```{r step6-year}
ggplot(df, aes(x = covariate1, y = yield_wweight, colour = factor(year))) +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE) +
  facet_wrap(~ year) +
  labs(title = "yield_wweight ~ covariate1 | year", colour = "year") +
  theme_bw() +
  theme(legend.position = "none")
```

## Coplot (Zuur style)

```{r step6-coplot}
coplot(yield_wweight ~ covariate1 | p_dist,
       data  = df,
       panel = function(x, y, ...) {
         points(x, y, pch = 19, col = "steelblue", cex = 0.7, ...)
         abline(lm(y ~ x), col = "red", lwd = 2)
       })
```

---

# Step 7 — Interactions {.tabset}

> **Zuur p. 11:** Non-parallel regression lines across groups indicate that
> an interaction term should be considered in the model.

## covariate1 × p_dist

```{r step7-pdist}
ggplot(df, aes(x = covariate1, y = yield_wweight,
               colour = p_dist, fill = p_dist, group = p_dist)) +
  geom_smooth(method = "lm", alpha = 0.15) +
  labs(title  = "Interaction check: yield_wweight ~ covariate1 × p_dist",
       colour = "p_dist", fill = "p_dist") +
  theme_bw()
```

## covariate1 × year

```{r step7-year}
ggplot(df, aes(x = covariate1, y = yield_wweight,
               colour = factor(year), fill = factor(year), group = factor(year))) +
  geom_smooth(method = "lm", alpha = 0.15) +
  labs(title  = "Interaction check: yield_wweight ~ covariate1 × year",
       colour = "year", fill = "year") +
  theme_bw()
```

---

# Step 8 — Independence of Observations {.tabset}

> **Zuur p. 12:** Nested sampling (block > id) and repeated measures across
> years both induce non-independence. ACF on naive residuals reveals temporal
> autocorrelation; residual plots by group reveal spatial clustering.

## ACF — temporal autocorrelation

```{r step8-acf}
m_naive <- lm(yield_wweight ~ covariate1 + p_dist, data = df)

acf(resid(m_naive),
    main = "ACF of naive residuals — temporal autocorrelation",
    col  = "steelblue",
    lwd  = 2)
```

## Residuals by group

```{r step8-residuals}
df_resid <- df |> mutate(.resid = resid(m_naive))

p_block <- ggplot(df_resid, aes(x = factor(block), y = .resid)) +
  geom_boxplot(fill = "plum", alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "red") +
  labs(title = "Residuals by block", x = "block", y = "residual") +
  theme_bw()

p_pdist <- ggplot(df_resid, aes(x = p_dist, y = .resid)) +
  geom_boxplot(fill = "skyblue", alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "red") +
  labs(title = "Residuals by p_dist", x = "p_dist", y = NULL) +
  theme_bw()

p_block + p_pdist
```

## Variance components (LMM)

```{r step8-lmm}
# block/id = nested random intercepts; year = repeated-measures grouping
m_lmm <- lmer(
  yield_wweight ~ p_dist + (1 | block/id) + (1 | year),
  data = df,
  REML = TRUE
)

summary(m_lmm)

# Proportion of total variance explained by each random effect
as.data.frame(VarCorr(m_lmm)) |>
  select(grp, vcov) |>
  mutate(pct_var = round(vcov / sum(vcov) * 100, 1)) |>
  arrange(desc(pct_var))
```

---

*Reference: Zuur AF, Ieno EN, Elphick CS (2010) A protocol for data exploration
to avoid common statistical problems. Methods in Ecology and Evolution 1: 3–14.
<https://doi.org/10.1111/j.2041-210X.2009.00001.x>*
